### 2011-06-17 uicthesi/uictest.tex
Bare structure, including uicthesi class (color modified) and using CC BY-NC-ND license.

% UICTEST.TEX
% This is a test file for my port of UICTHESI
% to LaTeX 2e. This is based in part on UCTHESIS.
%

\documentclass{uicthesi}

\begin{document}

% Declarations for Front Matter

\title{The Elements of Theses}
\author{Perry H. Disdainful}
\pdegrees{B.A. (University of Northern South Dakota at Hoople) 1978\\
M.S. (Ed's School of Quantum Mechanics and Muffler Repair Shop) 1989}
\degree{Doctor of Philosophy in Aboriginal Basketry}

\maketitle
\copyrightpage

\dedication
{\null\vfil
{\large
\begin{center}
To myself,\\\vspace{12pt}
Perry H. Disdainful,\\\vspace{12pt}
the only person worthy of my company.
\end{center}}
\vfil\null}

\acknowledgements
{I want to ``thank'' my committee, without whose ridiculous demands, I
would have graduated so, so, very much faster.}

\preface
This preface is purely optional at UIC.

\tableofcontents
\listoftables
\listoffigures
\listofabbreviations
\begin{list}
{}
{\setlength
   {\labelwidth}{1in}
    \setlength{\leftmargin}{1.5in}
    \setlength{\labelsep}{.5in}
    \setlength{\rightmargin}{\leftmargin}}
\item[AMS\hfill] American Mathematical Society
\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
\item[TUG\hfill] \TeX\ Users Group
\item[UIC\hfill] University of Illinois at Chicago
\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
\end{list}

\summary
A summary is required.

\chapter{Introduction}

Every dissertation should have an introduction.  You might not realize
it, but the introduction should introduce the concepts, backgrouand,
and goals of the dissertation.

\section{A Section}

Most chapters have sections. Neat huh?

\subsection{A Subsection}

Some chapters have subsections too! Cool!

\begin{table}
\begin{center}
\begin{tabular}{|l|r|}
  \hline 
Title & Author \\
\hline
War And Peace & Leo Tolstoy \\
The Great Gatsby & F. Scott Fitzgerald \\ \hline
\end{tabular}
\caption{A normalsize table.}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{A small table.}
\begin{tabular}{|l|r|}
  \hline 
Title & Author \\
\hline
War And Peace & Leo Tolstoy \\
The Great Gatsby & F. Scott Fitzgerald \\ \hline
\end{tabular}
\end{center}
\end{table}

\chapter{Previous Work}

Some other research was once performed.

\begin{figure}
\caption{A first figure.}
\end{figure}

\begin{figure}
\caption{A second figure.}
\end{figure}

\chapter{Conclusion}

A conclusion is a good thing to have. Hopefully it is a good one!

\appendices
\newpage
\appendix

\chapter{Some Ancillary Stuff}

Ancillary material should be put in appendices.

\chapter{Some More Ancillary Stuff}

Here is yet another appendix! Wahoo!

\nocite{*}
\bibformb
\bibliography{uictest}
\newpage
\vita
This is where the vita goes.  Its organization is left as an exercise.

\end{document}