### 2011-06-24 ms_model.tex

\section{Model of Michalik and Sipe}

This section introduces the Analytic Gaussian model of Michalik and Sipe.\cite{michalik_analytic_2006}

For a classical set of $N$ particles (labelled $i$), distributed 6D space (3 position, 3 momentum) by 
\begin{gather}
  f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = 
  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \\
  \text{which satisfies} \quad \int d^{3}x\,d^{3}p \; f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = N
\end{gather}
of mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
\begin{gather}
  \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j})
\end{gather}