### 2011-09-23 external_forces.tex
Work towards mathematical model

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\section{Extension to Include External Forces}

General form of (possibly multiple) external force ($\vec{F}_{ext}$) contributions
\begin{equation}
  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{force}_{ij} + \sum K^{ext}_{ij}
\end{equation}
in the basis
\begin{equation}
  U_{i} = \lbrace x, p_x, y, p_y, z, p_z \rbrace
\end{equation}
and
\begin{equation}
  K^{ext}_{ij} = \frac{1}{N} \iint U_i U_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
\end{equation}

For example, if $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 
\begin{equation}
  K^{ext} = M \cdot 
  \begin{pmatrix}
    K^{lens}_{\smallT} & 0 & 0 \\
    0 & K^{lens}_{\smallT} & 0 \\
    0 & 0 & 0
  \end{pmatrix}
  \qquad \text{where} \qquad
  K^{ext} = 
  \begin{pmatrix}
    0 & \sigma_{\smallT} \\
    \sigma_{\smallT} & 2 \gamma_{\smallT}
  \end{pmatrix}
\end{equation}
yielding an additional term $M_{\alpha} \sigma_{\alpha}$ which added to original $\frac{\partial \gamma_i}{\partial t}$ equation.