### 2011-09-29 inc/initial_conditions/initial_conditions.tex

\subsection{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}