### 2011-10-03 inc/initial_conditions/initial_conditions.tex
added CC notice to initial_conditions

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\subsection{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}