### 2012-12-12 inc/extension/implementation/Cooke.tex
added comment on Physics::UEMColumn and example script

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{}!/usr/bin/env perl}

\PY{k}{use} \PY{n}{strict}\PY{p}{;}
\PY{k}{use} \PY{n}{warnings}\PY{p}{;}

\PY{k}{use} \PY{n+nn}{Physics::UEMColumn} \PY{n}{alias} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{':standard'}\PY{p}{;}
\PY{k}{use} \PY{n+nn}{Physics::UEMColumn::Auxiliary} \PY{l+s}{':materials'}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser} \PY{o}{=} \PY{n}{Laser}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{width}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 mm'}\PY{p}{,}
  \PY{n}{duration} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 ps'}\PY{p}{,}
  \PY{n}{energy}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'4.75 eV'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc} \PY{o}{=} \PY{n}{DCAccelerator}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{length}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'20 mm'}\PY{p}{,}
  \PY{n}{voltage} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'20 kilovolts'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column} \PY{o}{=} \PY{n}{Column}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{length}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'400 mm'}\PY{p}{,} 
  \PY{n}{laser}        \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser}\PY{p}{,}
  \PY{n}{accelerator}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc}\PY{p}{,}
  \PY{n}{photocathode} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Photocathode}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}\PY{n}{Ta}\PY{p}{)}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n+nn}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{column}      \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column}\PY{p}{,}
  \PY{n}{number}      \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
  \PY{n}{debug}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf}       \PY{o}{=} \PY{l+m+mi}{20}\PY{p}{;} \PY{c+c1}{\PYZsh{}cm}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens} \PY{o}{=} \PY{l+s}{'1 in'}\PY{p}{;}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}  \PY{o}{=} \PY{l+m+mi}{5}\PY{p}{;} \PY{c+c1}{\PYZsh{}cm}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}    \PY{o}{=} \PY{l+m+mf}{33e-13}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens1} \PY{o}{=} \PY{n}{MagneticLens}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{-} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}\PY{p}{)} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens}\PY{p}{,}
  \PY{n}{strength} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens2} \PY{o}{=} \PY{n}{MagneticLens}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{+} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}\PY{p}{)} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens}\PY{p}{,}
  \PY{n}{strength} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens1}\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens2}\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{rf\PYZus{}cav} \PY{o}{=} \PY{n}{RFCavity}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'2 cm'}\PY{p}{,}
  \PY{n}{strength}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'230 kilovolts / m'}\PY{p}{,}
  \PY{n}{frequency} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'3 gigahertz'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{rf\PYZus{}cav}\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{propagate}\PY{p}{;}
\end{Verbatim}