### 2013-01-10 inc/extension/initial_conditions/bin_model_figure.tex
Implemented bin model result plots in the paper

\begin{figure}
  \centerline {
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Acc}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Vmax}
      \end{tikzpicture}
    }
    \\
  }
  \centerline{
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Equal}
      \end{tikzpicture}
    }
  }
\end{figure}