### 2013-01-30 inc/hardware/deflector.tex
added deflector content and schematic

\section{Deflection Plates}

In order to redirect the beam down the column, in conjunction with the accelerator alignment mechanisms, custom large area deflector plates were employed.
Four plates are arranged on a hollow tube with a square profile, one on each side.
One of each pair is held at ground while the other has a variable potential.
By varying each potential the beam may be turned in flight.

\begin{figure}
  \centering
  \include{deflector_schematic}
  \caption{
    Schematic diagram of one pair of deflection plates.
    The blue ray represents the path of the pulse as a whole (not to be confused with the plot of a lens).
  }
  \label{fig:deflector_schematic}
\end{figure}

A schematic description of such a system is shown in \ref{fig:deflector_schematic}.
To estimate the required voltage on the plates $V$, consider the applied impulse provided to the beam by the plate's electric field
\begin{gather}
  F \Delta t = m \Delta v \\
  \frac{ e V }{ d } \frac{ L }{ v_z } = m_e \Delta v_x
\end{gather}
where (as seen in the figure) $L$ is the length of the plate (in the direction of travel $\hat{z}$) and $d$ is the separation of the plates.
Then in the small angle approximation, the angle that the plate may correct for is approximately
\begin{equation}
  \theta = \frac{ \Delta v_x }{ v_z } = \frac{ e }{ m_e } \frac{ L }{ d } \frac{ V }{ v_z^2 }
\end{equation}
Using this result, when $L \sim d$ and $v_z = c/3$ a voltage of 1kV will deflect the beam approximately one degree.
These plates therefore are used as a fine adjustment, as opposed to the coarse adjustment provided by aligning the acceleration gap.