### 2013-02-15 inc/extension/initial_conditions/bin_model_figure.tex
comment on bin model

\begin{figure}
  \centerline {
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Acc}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Vmax}
      \end{tikzpicture}
    }
    \\
  }
  \centerline{
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Equal}
      \end{tikzpicture}
    }
  }
  \caption[Results of bin model for computing initial conditions]{Comparison of the results of the bin model and AG model allowed functional forms for three different input parameters cases.}
  \label{fig:bin_model}
\end{figure}