### 2013-02-25 inc/extension/initial_conditions/bin_model_figure.tex
more improvements to initial_conditions section

\begin{figure}
  \centerline {
    \subfloat[][]{
      \label{fig:bin_model_acc}
      \begin{tikzpicture}
        \input{Acc}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \label{fig:bin_model_vmax}
      \begin{tikzpicture}
        \input{Vmax}
      \end{tikzpicture}
    }
    \\
  }
  \centerline{
    \subfloat[][]{
      \label{fig:bin_model_equal}
      \begin{tikzpicture}
        \input{Equal}
      \end{tikzpicture}
    }
  }
  \caption[Results of bin model for computing initial conditions]{
    Comparison of the results of the bin model and AG model allowed functional forms for three different input parameters cases.
    Compared are: \subref{fig:bin_model_acc} the DC acceleration dominated case, \subref{fig:bin_model_vmax} the photoemission energy dominated case, and \subref{fig:bin_model_equal} an intermediate case.
    For $\sigma_z$ (red) and $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
    Unfortunately, the model only allows for a constant $\eta_z$ (green) and in all but the extreme acceleration dominated case \subref{fig:bin_model_acc} the model cannot well represent the simulated data.
  }
  \label{fig:bin_model}
\end{figure}