### 2013-03-01 main.tex
start merging andreas' comments on initial conditions section

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\documentclass{uicthesi}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{esint}
\usepackage{hyperref}
\usepackage{url}
\usepackage{ifthen}
\usepackage{pgffor}

%\usepackage{euler}
%\usepackage{breqn}

\usepackage[version=3]{mhchem}

\usepackage{import}
\usepackage{subfig}

\usepackage{graphicx}
\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.png,.pdf,.jpg}
\input{mytikz}
\setboolean{inputdata}{true}
\usepackage{fancyvrb}
\input{pygments}

%% My convenience fuctions
\providecommand{\vecprime}[1]{\vec{#1}\,^{\prime}}

\providecommand{\smallT}{\ensuremath{{\scriptscriptstyle T}}}
\providecommand{\smallD}{\ensuremath{{\scriptscriptstyle D}}}
\providecommand{\smallF}[0]{ { \scriptscriptstyle F } }
%\providecommand{\smallZ}{\ensuremath{{\scriptscriptstyle Z}}}

\providecommand{\punct}[1]{\ensuremath{\, \text{#1}}}

\providecommand{\abbr}[2]{\item[#1\hfill] #2}

\usepackage{color}
\providecommand{\fixme}[1]{\textcolor{red}{#1}}

%% Special handling for using biblatex vs bibtex ==
\newboolean{usebiblatex}
\setboolean{usebiblatex}{false}

\ifthenelse{\boolean{usebiblatex}}{
  \usepackage{biblatex}
  %\usepackage[backend=biber]{biblatex-chicago}
  \addbibresource{bibliography.bib}
  \addbibresource{biblaser.bib}
  \addbibresource{bibeste.bib}
}{}
%% ==

\title{Title}
\author{Joel Berger}
\pdegrees{of the\\University of Illinois at Chicago}
\degree{Doctor of Philosophy in Physics}

\begin{document}
\maketitle

%\copyrightpage
\newpage
\begin{figure}
  \centering
  \includegraphics{by-nc-nd}
\end{figure}
This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
To view a copy of this license, visit \url{http://creativecommons.org/licenses/by-nc-nd/3.0/us/} or send a letter to
\begin{verbatim}
Creative Commons
444 Castro Street, Suite 900
Mountain View, CA, 94041, USA.
\end{verbatim}

\dedication
The dedication.

\acknowledgment
The acknowledgment.
\initials{JAB}

\preface
The preface.

\tableofcontents
\listoftables
\listoffigures

\input{abbreviations}

\summary
A summary is required.

\chapter{Introduction}

\chapter{Review of Foundational Modeling Efforts} \label{chap:previous_models}

  \subimport{inc/ms_model/}{ms_model}

\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling} \label{chap:extension}

Modeling a system as complex as an Ultrafast Electron Microscope is an enourmous task.
In Chapter \ref{chap:previous_models}, I have highlighted some previous efforts at modeling electron pulse evolution.
While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.
In the following chapter, I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and is computationally fast enough to be used to design and optimize a column.

  \subimport{inc/extension/initial_conditions/}{initial_conditions}

  \subimport{inc/extension/liouvilles_theorem/}{liouvilles_theorem}

  %TODO refer from here to Free expansion results

  \subimport{inc/extension/external_forces/}{external_forces}

  \subimport{inc/gun/}{gun_model}

  \subimport{inc/extension/implementation/}{implementation}

\chapter{Model Results, Predictions and Implications}

\section{Free Expansion Propagation}

  \subimport{inc/model_results/shapes/}{shapes}

  \subimport{inc/model_results/spacecharge/}{spacecharge}

\section{Propagtion under External Forces}

  \subimport{inc/model_results/lens_charge/}{lens_charge}

\chapter{Prototype Instrument}

  \begin{figure}
    \centering
    \includegraphics{inc/hardware/column.jpg}
    \caption[Picture of the prototype UEM column at UIC]{
      Picture of the majority of the prototype UEM column at UIC.
      The main vacuum chamber is laid sideways, with the high voltage (gun) at the left and the imaging system on the right.
      The magnetic levitated turbo pump hangs below the column, a scroll pump used for backing is not pictured.
      The laser system is in the background.
      All of the above is mounted on a custom shaped active vibration-canceling optical table.
      The electronic control systems are not pictured.
    }
    \label{fig:column-pic}
  \end{figure}

  \subimport{inc/laser/}{laser}

  \subimport{inc/hardware/}{column}

  \subimport{inc/gun/}{gun_design}

  \subimport{inc/magnetic_lenses/}{maglens}

  \subimport{inc/hardware/}{deflector}

\chapter{Photocathode Engineering}

  \subimport{inc/photocathode/setup/}{setup}

\section{Flat Metal Photocathodes}

  \subimport{inc/photocathode/plasmon/}{plasmon}

  \subimport{inc/photocathode/este/}{este}

\chapter{Experimental Results}

%uncomment when multiple appedicies are used 
%\appendices
\appendix
  \subimport{inc/extension/implementation/}{appendix-script}

\newpage

%% Special handling for using biblatex vs bibtex ==
\ifthenelse{\boolean{usebiblatex}}{
  \printbibliography
}{
  \bibformc
  \bibliography{bibliography,biblaser,bibeste}
}
%% ==

%\vita
%This is where the vita goes.  Its organization is left as an exercise.
%Hint: see the list of abbreviations.
\end{document}