### 2013-03-15 inc/model_results/spacecharge/spacecharge.tex
expanded the spacecharge propagation section in model_results.
Added more realistic section which uses an accelerator too, though perhaps this should be moved to the model results with extension section, or maybe I should drop the distinction. hmmmm.

\subsection{Effect of Space Charge} \label{sec:free_spacecharge}

A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
Much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements.
Simplistically stated, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
On the other hand, space-charge effects are expected to dominate at large pulse charge densities.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_noacc}
  \end{tikzpicture}
  \caption[Simplistic free-space pulse evolution vs charge density]{
    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
    The pulse is initially 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
    It then traverses a 15cm drift region upon which the final length and width are recorded.
  }
  \label{fig:spacecharge_noacc}
\end{figure}

This trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) as a function of the number of electrons in the pulse after a propagation distance of 15cm.
The pulses are not accelerated but rather start with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
The simulations, clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
In the absence of space-charge effects, the presented prolate pulse shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.

%TODO highlight the following in the experimental section
%In other words, the spatial and temporal dispersive effects due to the nonzero value of $\Delta E$ are expected to dominate the space-time dynamics of the electron pulse during its propagation
%in our investigations because less than 10 5 electrons/pulse
%are generated.

As the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
The model predicts that space-charge effects produce a spherical charge distribution of approximately five times larger width at $10^8$ electrons/pulse for these conditions.
As noted in Section \ref{sec:initial_shapes}, pulses under the influence of the internal space-charge field expand preferentially in the dimension that was smallest initially; this causes the pulse to initially trend towards being spherical.

\ref{fig:spacecharge_noacc} is specifically created to isolate the behavior of and ``ideal'' pulse which is independent of its creation environment.
Of course in a more realistic system, the electron pulse is generated from a photocathode and then accelerated.
These forces play an interesting role in the system.

Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
The results of this simulation are presented in \ref{fig:spacecharge_acc}.
\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_acc}
  \end{tikzpicture}
  \caption[More realistic free-space pulse evolution vs charge density]{
    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy.
    It then traverses a 15cm column, a 20mm acceleration region followed by drift region.
  }
  \label{fig:spacecharge_acc}
\end{figure}