### 2013-03-21 inc/gun/gun_design.tex
tweak gun design intro now that design considerations presents the argument for large area emission

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\section{Accelerator Design} \label{sec:gun_design}

The acceleration region of our prototype UEM system has several requirement which are unique in comparison to standard electron guns.
As discussed in \ref{chap:considerations} notable requirements include the need to handle a large photoemission area and the inability to employ a pin-hole to selectively clean the electron beam.
This concern leads to several additional criteria.
The system must first be able to accept the large laser spot size needed to generate the electron beam.
Then, in order to accomodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to beam expansion during acceleration.
Also, care must be taken to ensure a relatively uniform electric field for acceleration of the pulse.
While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a wider proportion of the accelerator than the thin beams used in conventional TEM.

\begin{figure}
  \centering
  \centerline{\includegraphics{cathode.jpg} \includegraphics{anode.jpg}}
  \caption[Pictures of the employed Togawa-inspired custom cathode-anode pair]{
    Pictured are the cathode (left) and anode (right) of the accelerator built at UIC from a design by Togawa \protect\textit{et al.} \protect\cite{togawa_ceb6_2007}.
    The photocathode is visible in the cathode Wehnelt aperture.
    The anode is viewed as installed in the column, through the opened port normally occupied by the cathode and high voltage feedthrough.
    The white material seen in both images helps to prevent electrical arcing under high voltage.
  }
  \label{fig:togawa-pic}
\end{figure}

These requirements lead to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
The schematic shows that the aperture concerns are easily satisfied.
From the field equipotentials, it is clear that this design has a consistently flat electric field for large displacements off the central axis.

\begin{figure}
  \centering
  \input{gunfield}
  \caption[Schematic of the employed Togawa-inspired custom cathode-anode pair]{
    Schematic of the employed Togawa-inspired cathode ($-V$) and anode (0V) pair.
    The emission source is inserted behind the cathode Wehnelt.
    Also shown are simulated equipotential lines generated by FEM modeling.
    The separation of the cathode and anode is adjustable to allow a greater range of input laser angles if needed. 
  }
  \label{fig:gun-field}
\end{figure}

Of course, as derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.
This field will cause a divergence of the beam.
Future work will include investigating the designs of Butler %TODO reference
which may be able to build in an intrinsic lensing to compensate for this effect.

%TODO comment on alignment
%TODO comment on high voltage
%TODO comment on EV parts etc

Re-alignment of cathode and anode is sometimes necessary to direct the emitted beam down the center axis of the column.
To facilitate this, the prototype column has the photocathode and Wehnelt mounted on the end of the high voltage feedthrough, which is in turn mounted on a two axis alignment mechanism.
Although this mechanism, called a ``port aligner,'' is typically used to couple two non-colinear vacuum ports, for use in this instrument it has been modified to be able to manually adjust the aligner during use.
In contrast to this flexibility, the anode is mounted to column using Kimball Physics's ``eV Parts'' mounting systems.
These strictly manufactured mounting hardwares ensure that the anode is properly aligned with the remainder of the microscope column.
Thus, by iteratively repositioning the input laser and then the gun alignment, the beam may be walked to pass as near as possible down the axis of the column.

\begin{figure}
  \centering
  \includegraphics{aligner.jpg}
  \caption[Picture of the high-voltage side of the prototype UEM]{
    The high voltage passthrough (left) is mounted on the main chamber (right) by a ``port aligner'' which has been modified to allow coarse alignment of the cathode (mounted to the high voltage) and anode (mounted in the chamber).
  }
  \label{fig:aligner-pic}
\end{figure}