### 2013-03-26 inc/model_results/lens_charge/lens_charge_plot.tex
lens_charge is now all four permutations of long/short oblate/prolate

%TODO fix axes to be more consistent
%TODO add N=1e8 for prolate?

\begin{figure}
  \centering
  \centerline{
    \subfloat[][] {
      \label{fig:focus_long_prolate}
      \begin{tikzpicture}
        \input{lens_60mm_prolate_data}
        \node at (5.9,1.65) {$N=10^5$};
        \node at (4,3) {$N=10^6$};
        \node at (3,4) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] {
      \label{fig:focus_long_oblate}
      \begin{tikzpicture}
        \input{lens_60mm_oblate_data}
        \node at (4.2,2.3) {$N=10^5$};
        \node at (3.3,3.5) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \centerline{
    \subfloat[][] { 
      \label{fig:focus_short_prolate}
      \begin{tikzpicture}
        \input{lens_6mm_prolate_data}
        \node at (6.5,3) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] { 
      \label{fig:focus_short_oblate}
      \begin{tikzpicture}
        \input{lens_6mm_oblate_data}
        \node at (6.9,2.6) {$N=10^7$};
        \node at (6,2) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
    All plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
    The left plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are prolate shaped, initially HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
    The right plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are oblate shaped, initially HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.
    The top plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_long_oblate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_prolate} and \subref{fig:focus_short_oblate}) $ f = $ 6.0mm.
    A logarithmic scale is used for comparative clarity.
    Clearly the best performance is achieved for shorter focal lengths and prolate pulses even at higher initial charge densities.
  }
  \label{fig:focus_lens_charge}
\end{figure}