### 2013-03-27 inc/model_results/lens_charge/lens_charge.tex
fix oblate/prolate gaff (ooops)

\subsection{Space-charge Effects in Magnetic Lenses} \label{sec:mag_lens_charge}

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses.\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry.\cite{berger_dc_2009}
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity).

In the preceding sections it has already been established that oblate eccentricities are preferred for UEM in general.
When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.

\input{lens_charge_plot}

\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for oblate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
To make an effective comparison, highlighting the effect of space-charge alone, the pulses used are at 20kV velocity with $\Delta E = $ 0.5eV but are not simulated using an accelerator.
The length of the accelerator and the anode lensing would make useful comparison difficult; as such the simulation presented approximates a beam which has been perfectly collimated at the front-focal plane of the lens.

The single-electron ($ N = 1 $) focal spot sizes of $\sim$150 microns for the $ f = 60 \text{mm} $ lens and $\sim$15 microns for the shorter $ f = 6.0 \text{mm} $ lens show marked increase starting around $ N \approx 10^{ 5 } $ and $ N \approx 10^{ 6 }$, respectively, and worsening dramatically above these thresholds.
This order of magnitude difference is fundamentally related to the greater time of flight before the focus of the $ f = 60\text{mm} $ lens; it increases the impulse that the internal Coulomb force can exert on the pulse as its charge density is increased under focusing.
A consequent further effect is a shift in the focal position to $ z > f $.
%A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}).

%TODO? include simulations for different voltages to establish TOF argument?

Thus, the extended AG electron pulse propagation model predicts that higher fidelity focusing will be achieved with shorter focal length magnetic lenses, higher acceleration energy electrons (time of flight reduction), and pulses with $ \sigma_{ \smallT } \gg \sigma_{ z } $ at the entrance aperture of the lens.