### 2013-03-28 inc/model_results/compound/compound.tex
added compound element section (heading) and demonstration figure

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\subsection{Compound Element Systems} \label{sec:compound}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{compound_data}
  \end{tikzpicture}
  \caption[Example model of a compound system of electron-optical elements]{
    Example model of a compound system of electron-optical elements.
  }
  \label{fig:compound}
\end{figure}