### 2013-03-30 inc/photocathode/setup/fig1.tex
remove (essentially) unused \inputdata command, which was meant to ignore loading tikz data

\begin{tikzpicture}
  \fill [blue!20] 
    (1.3,0.95)
    -- ++(0.7,0)
    -- ++(0,7.5)
    -- ++(-0.7,0)
      node [black, above, pos=0.5, align=center] {Acceleration\\Region}
    -- cycle
  ;
  \fill [blue!20]
    (4.4,0.95)
    -- ++(2,0)
    -- ++(0,7.5)
    -- ++(-2,0)
      node [black, pos=0.5, above, align=center] {Magnetic\\Lenses}
    -- cycle
  ;
  \fill [blue!20] 
    (10.5,0.95)
    -- ++(0.7,0)
    -- ++(0,7.5)
    -- ++(-0.7,0)
      node [black, above, pos=0.5] {Detector}
    -- cycle
  ;
  \input{fig1data}
\end{tikzpicture}