### 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
added rf cavity comparison plot (like figs 5(a)/6(a) from JAP paper)

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\section{Space-charge Effects in RF Cavities} \label{sec:rf_cav_charge}

\input{rf_cav_plot}