### 2013-04-01 inc/model_results/rf_cav/rf_cav_plot.tex

\begin{figure}
  \centerline {
    \subfloat[][]{
      \label{fig:rf_cav_N1}
      \begin{tikzpicture}
        \input{rf_cav_N1}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \label{fig:rf_cav_N1e5}
      \begin{tikzpicture}
        \input{rf_cav_N1e5}
      \end{tikzpicture}
    }
  }
  \caption[Short]{
    Long
  }
  \label{fig:rf_cav_num}
\end{figure}