### 2013-04-01 inc/model_results/rf_cav/rf_cav_plot.tex
tweaks to the rf_cav model_results section, including a figure caption

\begin{figure}
  \centerline {
    \subfloat[][]{
      \label{fig:rf_cav_N1}
      \begin{tikzpicture}
        \input{rf_cav_N1}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \label{fig:rf_cav_N1e5}
      \begin{tikzpicture}
        \input{rf_cav_N1e5}
      \end{tikzpicture}
    }
  }
  \caption[Dynamics of the pulse length and width under the influence of an RF cavity]{
    Dynamics of the pulse length (green) and width (red) under the influence of an RF cavity placed at $z' = 0$.
    The employed idealized pulses start at $z'= -100$ cm with velocity $v_{\smallzero} = \sim 84 \times 10^{6}$ m/s (20kV) and initial HW1/eM width of 100 $\mu$m and length $\sim 8.4 \mu$m ($v_{\smallzero} \cdot $ 100fs).
    \subref{fig:rf_cav_N1} has $N=1$ electron and demonstrates one-to-one imaging of the initial pulse length at the focus at $ z' = 100 $ cm.
    \subref{fig:rf_cav_N1e5} has $ N = 1 \times 10^5 $ electrons, whose added broadening allows for the factor of two reduction in pulse length at the focus ($\sim 48$ fs).
  }
  \label{fig:rf_cav_num}
\end{figure}