### 2013-04-02 inc/model_results/compound/compound_plot.tex

\begin{tikzpicture}
  \draw [dashed,fill=blue!20]
    (1.5,8.375)
    -- ++(0,-7.4)
    -- ++(0.6,0)
    -- ++(0,7.4)
    -- ++(-0.6,0)
      node [pos=0.5,rotate=30,above right] {Acceleration Region}
    -- cycle
  ;
  \draw [dashed]
    (5.73,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Magnetic Lens}
  ;
  \draw [dashed]
    (7.16,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {RF Cavity}
  ;
  \draw [dashed]
    (8.55,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Magnetic Lens}
  ;
  \draw [dashed]
    (9.975,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Focus}
  ;
  \input{compound_data}
\end{tikzpicture}