### 2013-04-08 inc/photocathode/future.tex
working on cleaning up the splitting out of parts of este section

\section{Future Prospects} \label{sec:photocathode_future}

A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
 and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}