### 2013-04-08 inc/photocathode/future.tex

\section{Future Prospects} \label{sec:photocathode_future}

A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m$ \cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_{\smallT}$.
 and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}