### 2013-04-10 inc/hardware/deflector/deflector.tex
merged remaining corrections from Carolyn

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\section{Deflection Plates}

In order to direct the beam down the column, custom large area deflector plates are employed in conjunction with the gross accelerator alignment mechanism described earlier.
Four plates are arranged on a hollow tube with a square profile, one on each side.
One of each pair is held at ground while the other has a variable potential.
By varying each potential the beam may be turned in flight.

\begin{figure}
  \centering
  \include{deflector_schematic}
  \caption[Schematic diagram of one pair of deflection plates]{
    Schematic diagram of one pair of deflection plates.
    The blue ray represents the path of the pulse as a whole (not to be confused with the plot of a lens).
  }
  \label{fig:deflector_schematic}
\end{figure}

A schematic description of such a system is shown in \ref{fig:deflector_schematic}.
To estimate the required voltage on the plates $V$ needed to correct for a lateral velocity $v_x$, consider the applied impulse provided to the beam by the plate's force $F$ of the electric field,
\begin{gather}
  F \Delta t = m \Delta v \\
  \frac{ e V }{ d } \frac{ L }{ v_z } = m \Delta v_x \,\text{,}
\end{gather}
where (as seen in the figure) $L$ is the length of the plate (in the direction of travel $\hat{z}$) and $d$ is the separation of the plates.
If the beam has a velocity $v_z$ down the axis of the column, then in the small angle approximation, the angle that the plate may correct for is approximately
\begin{equation}
  \theta = \frac{ \Delta v_x }{ v_z } = \frac{ e }{ m } \frac{ L }{ d } \frac{ V }{ v_z^2 } \,\text{.}
\end{equation}
Using this result, when $L \sim d$ and $v_z = c/3$ a voltage of 1kV will deflect the beam approximately one degree.
These plates are, therefore, useful as a fine beam adjustment, as contrast to the coarse adjustment provided by manually aligning the acceleration gap.

\begin{figure}
  \centering
  \includegraphics{deflector_plates.jpg}
  \caption[Picture of custom deflector plates]{
    Picture of the custom built deflector plates for use in the UEM at UIC.
    A pair of vertical and horizontal plates are mounted in a non-conductive housing.
    Each plate has a pin for connecting to voltage or ground.
    Clearance drilled holes allow mounting onto the ``eV Parts'' cage inside the chamber.
  }
  \label{fig:deflector-plates-pic}
\end{figure}