### 2013-04-11 inc/extension/introduction.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter

In Chapter \ref{chap:ms_model}, I have introduced the AG model, originally presented by Michalik and Sipe~\cite{michalik_analytic_2006}.
Although it is a very impressive model and a very useful starting point, because it cannot include the influence of external forces on the pulse dynamics, it does not provide the complete picture of the evolution of an electron pulse through the entire column.
Futher, it does not contain a set of initial conditions which reflect a realistic photoemission process.

In the following chapter, I will derive a set of initial conditions which should be sufficient to model single-photon photoemission from a metal, and which may serve as a basis for future implementations employing other photoemission processes.
I will then present a model which builds on and extends the AG model to include the influence of external forces on pulses.
I will also include the mathematical representations of magnetic lenses, TM$_{010}$ RF cavities, and DC accelerators.
With these computational tool it is possible to form one model which has the ability to simulate a UEM column.
Indeed I have implemeted this model in the Perl programming language, in a form generic enough to model many different columns and yet is computationally fast enough to be used to design and optimize a column.