### 2013-04-11 main.tex

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\documentclass{uicthesi}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{esint}
\usepackage{hyperref}
\usepackage{url}
\usepackage{ifthen}
\usepackage{pgffor}

%\usepackage{euler}
%\usepackage{breqn}

\usepackage[version=3]{mhchem}

\usepackage{import}
\usepackage{subfig}

\usepackage{graphicx}
\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.png,.pdf,.jpg}

\usepackage[noexternal]{mytikz}

\usepackage{fancyvrb}
\input{pygments}

%% My convenience fuctions
\providecommand{\vecprime}[1]{\vec{#1}\,^{\prime}}

\providecommand{\smallT}{\ensuremath{{\scriptscriptstyle T}}}
\providecommand{\smallD}{\ensuremath{{\scriptscriptstyle D}}}
\providecommand{\smallzero}{\ensuremath{{\scriptscriptstyle 0}}}
\providecommand{\smallF}[0]{ { \scriptscriptstyle F } }
%\providecommand{\smallZ}{\ensuremath{{\scriptscriptstyle Z}}}

\providecommand{\punct}[1]{\ensuremath{\, \text{#1}}}

\providecommand{\abbr}[2]{\item[#1\hfill] #2}

\usepackage{color}
\providecommand{\fixme}[1]{\textcolor{red}{#1}}

%% Special handling for using biblatex vs bibtex ==
\newboolean{usebiblatex}
\setboolean{usebiblatex}{false}

\ifthenelse{\boolean{usebiblatex}}{
  \usepackage{biblatex}
  %\usepackage[backend=biber]{biblatex-chicago}
  \addbibresource{bibliography.bib}
  \addbibresource{biblaser.bib}
  \addbibresource{bibeste.bib}
}{}
%% ==

\title{Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling}
\author{Joel Berger}
\pdegrees{of the\\University of Illinois at Chicago}
\degree{Doctor of Philosophy in Physics}

\begin{document}
\maketitle

%\copyrightpage
\newpage
\begin{figure}
  \centering
  \includegraphics{by-nc-nd}
\end{figure}
This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
To view a copy of this license, visit \url{http://creativecommons.org/licenses/by-nc-nd/3.0/us/} or send a letter to
\begin{verbatim}
Creative Commons
444 Castro Street, Suite 900
Mountain View, CA, 94041, USA.
\end{verbatim}

\dedication
The dedication.

\acknowledgment
The acknowledgment.
\initials{JAB}

\preface
The preface.

\tableofcontents
%\listoftables
\listoffigures

\input{abbreviations}

\summary
A summary is required.

\chapter{Introduction}

  \subimport{inc/}{introduction}

\chapter{The Analytic Gaussian Model} \label{chap:ms_model}

  \subimport{inc/ms_model/}{ms_model}

\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling} \label{chap:extension}

  \subimport{inc/extension/}{introduction}

  \subimport{inc/extension/initial_conditions/}{initial_conditions}

  \subimport{inc/extension/liouvilles_theorem/}{liouvilles_theorem}

  \subimport{inc/extension/external_forces/}{formal}

  \subimport{inc/extension/external_forces/}{practical}

  \subimport{inc/gun/}{gun_model}

  \subimport{inc/extension/implementation/}{implementation}

\chapter{Model Results, Predictions and Implications} \label{chap:model_results}

  \subimport{inc/model_results/}{introduction}

  \subimport{inc/model_results/shapes/}{shapes}

  \subimport{inc/model_results/spacecharge/}{spacecharge}

  \subimport{inc/model_results/lens_charge/}{lens_charge}

  \subimport{inc/model_results/rf_cav/}{rf_cav}

  \subimport{inc/model_results/compound/}{compound}

\chapter{UEM Design Considerations} \label{chap:considerations}

  \subimport{inc/design_considerations/}{considerations}

\chapter{Prototype Instrument} \label{chap:prototype}

  \subimport{inc/hardware/}{introduction}

  \subimport{inc/hardware/laser/}{laser}

  \subimport{inc/hardware/}{column}

  \subimport{inc/gun/}{gun_design}

  \subimport{inc/hardware/magnetic_lenses/}{maglens}

  \subimport{inc/hardware/deflector/}{deflector}

\chapter{Photocathode Engineering} \label{chap:photocathode}

  \subimport{inc/photocathode/setup/}{setup}

  \subimport{inc/photocathode/single_photon/}{single_photon}

  \subimport{inc/photocathode/two_photon_thermionic/}{two_photon_thermionic}

  \subimport{inc/photocathode/plasmon/}{plasmon}

  \subimport{inc/photocathode/este/}{este}

  \subimport{inc/photocathode/}{future}

\chapter{Conclusions}

  \subimport{inc/}{conclusions}

%uncomment when multiple appedicies are used 
%\appendices
\appendix
  \subimport{inc/extension/implementation/}{appendix-script}

\newpage

%% Special handling for using biblatex vs bibtex ==
\ifthenelse{\boolean{usebiblatex}}{
  \printbibliography
}{
  \bibformc
  \bibliography{bibliography,biblaser,bibeste}
}
%% ==

%\vita
%This is where the vita goes.  Its organization is left as an exercise.
%Hint: see the list of abbreviations.
\end{document}