### 2013-04-12 inc/appendix/script.tex

\chapter{Example Simulation Script} \label{sec:script}

\begin{singlespace}
  \scriptsize
  \input{compound}
\end{singlespace}

\begin{singlespace}
  \scriptsize
  \input{spacecharge}
\end{singlespace}