### 2013-07-21 vita.tex
started vita section

\vita

\begin{center}
\begin{singlespace}

\vspace{0.4in}

\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}

{NAME:}
& Joel Adam Berger \\
\\
{EDUCATION:}
& B.S., Physics, University of Illinois at Chicago,\\
& Chicago, Illinois, 2005 \\
\\
{\Large TEACHING:}
& Something here\\

\end{tabular}

\end{singlespace}
\end{center}