### 2013-07-21 vita.tex
fix broken vita markup

\vita

\begin{center}
\begin{singlespace}

\vspace{0.4in}

%\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}
\begin{tabular}{lp{4.5in}}
NAME: & Joel Adam Berger \\
\\
EDUCATION: & B.S., Physics, University of Illinois at Chicago,\\
& Chicago, Illinois, 2005 \\
\\
TEACHING: & Something here\\
\end{tabular}

\end{singlespace}
\end{center}