### 2013-07-23 inc/copyright.tex
include copyright notice re optics express

\noindent \ref{fig:crystal}\subref{fig:crystal_coating}, \ref{fig:laser-thermal}, \ref{fig:laser-cavity}\subref{fig:laser_stability}, \ref{fig:laser-profile} and \ref{fig:laser-performance} were previously published in Optics Express. The full article is available at \url{http://www.opticsinfobase.org/oe/abstract.cfm?uri=oe-16-12-8629}. Their use of these figures is granted to the author under the permissive non-commercial terms of the copyright transfer agreement (\url{http://www.opticsinfobase.org/oe/submit/review/copyright_permissions.cfm}).